---
title: "Manegerial Segmentation"
author: "Abhishek Unni Krishnan"
date: "12/07/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sqldf)
```
*Problem Statement:Carry out a Segmentation in order to find out which group of customers should be targetted in order to increase the revenue*

*Part1: Reading and Formating Data*
```{r}
getwd()
setwd("C:\\Users\\abhit\\OneDrive\\Documents\\Uni\\Marketing analytics") 
#Reading the Data
Seg1 = read.delim(file = 'rfm_data.txt', header = FALSE, sep = '\t', dec = '.')
#looking at the imported data
View(Seg1)
colnames(Seg1) = c('customer_id', 'purchase_amount', 'date_of_purchase')
Seg1$date_of_purchase = as.Date(Seg1$date_of_purchase, "%Y-%m-%d")
Seg1$year_of_purchase = as.numeric(format(Seg1$date_of_purchase, "%Y"))
Seg1$days_since       = as.numeric(difftime(time1 = "2016-01-01",
                                            time2 = Seg1$date_of_purchase,
                                            units = "days"))

head(Seg1)
```
*Part2: Calculating Recency Frequency and Average Purchase Amount for all the data till the year 2015*
```{r}
# Compute recency, frequency, and Average purchase amount
customers_2015 = sqldf("SELECT customer_id,
                               MIN(days_since) AS 'recency',
                               MAX(days_since) AS 'first_purchase',
                               COUNT(*) AS 'frequency',
                               AVG(purchase_amount) AS 'amount'
                        FROM Seg1 GROUP BY 1")

#Creating segments based on our data for the year 2015
customers_2015$segment = "NA"
customers_2015$segment[which(customers_2015$recency > 365*3)] = "inactive"
customers_2015$segment[which(customers_2015$recency <= 365*3 & customers_2015$recency > 365*2)] = "cold"
customers_2015$segment[which(customers_2015$recency <= 365*2 & customers_2015$recency > 365*1)] = "warm"
customers_2015$segment[which(customers_2015$recency <= 365)] = "active"
customers_2015$segment[which(customers_2015$segment == "warm" & customers_2015$first_purchase <= 365*2)] = "new warm"
customers_2015$segment[which(customers_2015$segment == "warm" & customers_2015$amount < 100)] = "warm low value"
customers_2015$segment[which(customers_2015$segment == "warm" & customers_2015$amount >= 100)] = "warm high value"
customers_2015$segment[which(customers_2015$segment == "active" & customers_2015$first_purchase <= 365)] = "new active"
customers_2015$segment[which(customers_2015$segment == "active" & customers_2015$amount < 100)] = "active low value"
customers_2015$segment[which(customers_2015$segment == "active" & customers_2015$amount >= 100)] = "active high value"
table(customers_2015$segment)
aggregate(x = customers_2015[, 2:5], by = list(customers_2015$segment), mean)
```
*Part3: Calculating Recency Frequency and Average Purchase Amount for all the data till the year 2014*
```{r}
# Compute recency, frequency, and average purchase amount
customers_2014 = sqldf("SELECT customer_id,
                               MIN(days_since) - 365 AS 'recency',
                               MAX(days_since) - 365 AS 'first_purchase',
                               COUNT(*) AS 'frequency',
                               AVG(purchase_amount) AS 'amount'
                        FROM Seg1
                        WHERE days_since > 365
                        GROUP BY 1")


customers_2014$segment = "NA"
customers_2014$segment[which(customers_2014$recency > 365*3)] = "inactive"
customers_2014$segment[which(customers_2014$recency <= 365*3 & customers_2014$recency > 365*2)] = "cold"
customers_2014$segment[which(customers_2014$recency <= 365*2 & customers_2014$recency > 365*1)] = "warm"
customers_2014$segment[which(customers_2014$recency <= 365)] = "active"
customers_2014$segment[which(customers_2014$segment == "warm" & customers_2014$first_purchase <= 365*2)] = "new warm"
customers_2014$segment[which(customers_2014$segment == "warm" & customers_2014$amount < 100)] = "warm low value"
customers_2014$segment[which(customers_2014$segment == "warm" & customers_2014$amount >= 100)] = "warm high value"
customers_2014$segment[which(customers_2014$segment == "active" & customers_2014$first_purchase <= 365)] = "new active"
customers_2014$segment[which(customers_2014$segment == "active" & customers_2014$amount < 100)] = "active low value"
customers_2014$segment[which(customers_2014$segment == "active" & customers_2014$amount >= 100)] = "active high value"
table(customers_2014$segment)
```
*Part 4: Comparing distribution of the customer segment in the year 2014 and 2015*
```{r}
pie(table(customers_2014$segment), main="Segmentation in 2014") #2014

pie(table(customers_2015$segment), main="Segmentation in 2015") #2015
```




*Part 5: Calculating Average revenue per Segment for the year 2015*
```{r}
revenue_2015 = sqldf("SELECT customer_id, SUM(purchase_amount) AS 'revenue_2015'
                      FROM Seg1
                      WHERE year_of_purchase = 2015
                      GROUP BY customer_id")
View(revenue_2015)
summary(revenue_2015)
```
*Now Merging with the 2014 DATA in order to compare how the customers per Segment in the year 2014 spent in year 2015* 
```{r}
#merging average revenue of 2015 customer with original table
actual = merge(customers_2015, revenue_2015, all.x = TRUE)
actual$revenue_2015[is.na(actual$revenue_2015)] = 0
#avg revenue per customer per segment
aggregate(x = actual$revenue_2015, by = list(customers_2015$segment), mean)
#merging 2014 and 2015 data
forward = merge(customers_2014, revenue_2015, all.x = TRUE)
forward$revenue_2015[is.na(forward$revenue_2015)] = 0
View(forward)
```
*Part 6: Finding Average Revenue Per Segment in the year 2015*
```{r}
r = aggregate(x = forward$revenue_2015, by = list(customers_2014$segment), mean)
#arranging it in decreasing order
r = r[order(r$x, decreasing = TRUE), ]
print(r)
barplot(r$x, names.arg = r$Group.1)
```




*Part 7: Inference*


*Based on the results we got we can clearly see that our new active members tend to have a very low revenue in the following year so as a company we have to target these customers in order to keep them.This could be done by increasing a promotion or introduction of a points system in order to increase their average spend.*